<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{An R Markdown Vignette made with knitr}
-->

Statistics Finland (Tilastokeskus) R tools
===========

This [statfi](http://ropengov.github.com/statfi) R package to access
open data from [Statistics
Finland](http://www.stat.fi/tup/tilastotietokannat/index_fi.html). 

The available data sets include about 3000 data sets from [Statistics
Finland](http://www.stat.fi/org/lainsaadanto/avoin_data.html),
[Eurostat](http://pxweb2.stat.fi/Database/Eurostat/databasetree_fi.asp),
and other [international
statistics](http://pxweb2.stat.fi/Database/Kansainvalisen_tiedon_tietokanta/databasetree_fi.asp)). The
package is part of the [rOpenGov](http://ropengov.github.io) project.

## Installation

Release version for general use:

```{r install, eval=FALSE}
install.packages("statfi")
library(statfi)
```

Development version (potentially unstable):

```{r install2, eval=FALSE}
install.packages("devtools")
library(devtools)
install_github("statfi", "ropengov")
library(statfi)
```


## Listing of available data sets (Statistics Finland, Eurostat, International)

The listings of [open data sets from Statistics Finland (StatFi),
Eurostat and International
statistics](http://www.stat.fi/org/lainsaadanto/avoin_data.html) are
available for browsing in PCAxis, CSV and XML format:

 * StatFi [PCAxis](http://pxweb2.stat.fi/database/StatFin/databasetree_fi.asp) [CSV](http://pxweb2.stat.fi/database/StatFin/StatFin_rap_csv.csv) [XML](http://pxweb2.stat.fi/database/StatFin/StatFin_rap_xml.csv)

 * Eurostat [PCAxis](http://pxweb2.stat.fi/Database/Eurostat/databasetree_fi.asp) [CSV](http://pxweb2.stat.fi/database/StatFin/StatFin_rap.csv)

 * International statistics [PC Axis](http://pxweb2.stat.fi/Database/Kansainvalisen_tiedon_tietokanta/databasetree_fi.asp)

You can download these listings in R as follows:

```{r tilastokeskus-1, warning=FALSE, message=FALSE}
# Load the library
library(statfi)

# List open data files available from Statistics Finland
datasets.statfi <- list_statfi_files()

# List open data files available from Eurostat
datasets.eurostat <- list_eurostat_files()

# Investigate the first entry in StatFi data
print(datasets.statfi[1,])

# Descriptions of the first entries
head(datasets.statfi$DESCRIPTION)
```

## Retrieving the data

This example illustrates how to retrieve data from the StatFi
databases by defining the URL of the data set. For the listing of
available files and URLs, see above.

```{r tilastokeskus-2, warning=FALSE, message=FALSE}
library(statfi)

# Define URL (see list_statfi_files() for listing of available files)
url <- "http://pxweb2.stat.fi/Database/StatFin/tul/tvt/2009/120_tvt_2009_2011-02-18_tau_112_fi.px"

# Get the data
df <- get_statfi(url)
df[1:3,]
```

For further usage examples, see
[Louhos-blog](http://louhos.wordpress.com) and
[Datawiki](https://github.com/ropengov/statfi/wiki/Data).


## Licensing and Citations

### Statistics Finland data

Cite Statfi and link to
[http://www.statfi.fi](http://www.statfi.fi/). We are grateful to
Statistics Finland open data personnell for their generous support
during the development of this package.

### statfi R package

This work can be freely used, modified and distributed under the
[Two-clause FreeBSD
license](http://en.wikipedia.org/wiki/BSD\_licenses). Kindly cite the
R package as 'Leo Lahti, Juuso Parkkinen ja Joona Lehtomäki
(2010-2013). statfi R package. URL:
http://ropengov.github.io/statfi'.


### Session info


This vignette was created with

```{r sessioninfo, message=FALSE, warning=FALSE}
sessionInfo()
```
